\setcounter{table}{0}
\renewcommand{\thetable}{\thechapter.\arabic{table}}

\begin{longtable}{p{0.22\textwidth}p{0.12\textwidth}p{0.18\textwidth}p{0.34\textwidth}}
    \caption{Characteristics of Included Studies} \label{tab:included_studies}                                                                                                                                                                                                                                                                                                                                                          \\
    \toprule
    \textbf{Study}                                                                                                                     & \textbf{Country}  & \textbf{AI Technology}            & \textbf{Primary Goal}                                                                                                                                                                                                                  \\
    \midrule
    \endfirsthead

    \multicolumn{4}{c}{\tablename\ \thetable\ -- \textit{Continued from previous page}}                                                                                                                                                                                                                                                                                                                                                 \\
    \toprule
    \textbf{Study}                                                                                                                     & \textbf{Country}  & \textbf{AI Technology}            & \textbf{Primary Goal}                                                                                                                                                                                                                  \\
    \midrule
    \endhead

    \midrule
    \multicolumn{4}{r}{\textit{Continued on next page}}                                                                                                                                                                                                                                                                                                                                                                                 \\
    \endfoot

    \bottomrule
    \endlastfoot

    \citeauthor{balsaUsabilityIntelligentVirtual2020}, \citeyear{balsaUsabilityIntelligentVirtual2020}                                 & Portugal          & Knowledge-Based \& Rule-Based     & To promote behavior change and self-care, including healthy diet, in older people with Type 2 Diabetes through an intelligent virtual assistant.                                                                                       \\
    \addlinespace
    \citeauthor{chatterjeeDesigningInternetofThingsIoT2018}, \citeyear{chatterjeeDesigningInternetofThingsIoT2018}                     & USA               & Deep Learning                     & To change dietary behaviour and predict next-day blood glucose levels using daily activity, physiological, and nutritional data.                                                                                                       \\
    \addlinespace
    \citeauthor{chewADLDigitalToolkit2025}, \citeyear{chewADLDigitalToolkit2025}                                                       & Singapore         & Deep Learning                     & To monitor food intake, analyze nutritional content from meal photos, and promote healthier eating habits using an AI-driven tool.                                                                                                     \\
    \addlinespace
    \citeauthor{chopraDigiMateLeveragingLarge2024}, \citeyear{chopraDigiMateLeveragingLarge2024}                                       & USA               & Natural Language Processing (NLP) & To provide a personalized digital coach for the mobility challenged elderly population to provide personalized guidance for controlling obesity.                                                                                       \\
    \addlinespace
    \citeauthor{chunDeepLearningBased2024}, \citeyear{chunDeepLearningBased2024}                                                       & South Korea       & Deep Learning                     & To classify the freshness of fruits (apples, bananas, oranges) from smartphone images to support older adults in identifying rotten food.                                                                                              \\
    \addlinespace
    \citeauthor{dimartinoMalnutritionRiskAssessment2021}, \citeyear{dimartinoMalnutritionRiskAssessment2021}                           & Italy             & Machine Learning (General)        & To provide a Decision Support System for the early detection of malnutrition risk in frail older adults.                                                                                                                               \\
    \addlinespace
    \citeauthor{gautamFoodAssessmentModel2022}, \citeyear{gautamFoodAssessmentModel2022}                                               & India             & Deep Learning                     & To accurately classify and estimate the food region size of mixed/cooked food images, which can be used to obtain calories and nutrient information for Indian elderly persons.                                                        \\
    \addlinespace
    \citeauthor{grecoEmotionalFeaturesInteractions2021}, \citeyear{grecoEmotionalFeaturesInteractions2021}                             & Italy             & Natural Language Processing (NLP) & To assist elderly users in improving nutritional habits through a goal-directed conversation.                                                                                                                                          \\
    \addlinespace
    \citeauthor{haiSelfFeedingRobotElder2022}, \citeyear{haiSelfFeedingRobotElder2022}                                                 & Vietnam           & Machine Learning (General)        & To analyze a patient's nutritional consumption based on meal records from a self-feeding robot and provide diet recommendations.                                                                                                       \\
    \addlinespace
    \citeauthor{ioakeimidisCanBiteDetection2024}, \citeyear{ioakeimidisCanBiteDetection2024}                                           & Sweden            & Deep Learning                     & To automatically detect and count mouthfuls from video recordings of meals to monitor eating behavior in elderly people with and without Parkinson's disease.                                                                          \\
    \addlinespace
    \citeauthor{khattakSelfAdaptiveSystem2016}, \citeyear{khattakSelfAdaptiveSystem2016}                                               & UAE               & Knowledge-Based \& Rule-Based     & To provide personalized nutrition and activity recommendations for the elderly based on sensor-recognized daily activities and user profile.                                                                                           \\
    \addlinespace
    \citeauthor{kobayashiFrailtyMeterHealthy2021}, \citeyear{kobayashiFrailtyMeterHealthy2021}                                         & Japan             & Deep Learning                     & To automatically calculate the Dietary Variety Score from food photos to assess the nutritional component of frailty.                                                                                                                  \\
    \addlinespace
    \citeauthor{kyritsisAssessmentRealLife2021}, \citeyear{kyritsisAssessmentRealLife2021}                                             & Greece            & Deep Learning                     & To objectively classify individuals as Parkinson's Disease patients or Healthy Controls based on eating behavior profiles derived from smartwatch sensor data.                                                                         \\
    \addlinespace
    \citeauthor{liKGSCSSmartCare2024}, \citeyear{liKGSCSSmartCare2024}                                                                 & China             & Knowledge-Based \& Rule-Based     & To generate a 'diet information' section within a comprehensive care report, including risk identification and care recommendations, for elderly individuals with chronic diseases.                                                    \\
    \addlinespace
    \citeauthor{obaidAutomaticFoodintakeMonitoring2018}, \citeyear{obaidAutomaticFoodintakeMonitoring2018}                             & USA               & Computer Vision (Non-DL)          & To monitor food-intake for patients with Alzheimer's by tracking hand-to-mouth movements from video.                                                                                                                                   \\
    \addlinespace
    \citeauthor{papathanailEvaluationNovelArtificial2021}, \citeyear{papathanailEvaluationNovelArtificial2021}                         & Switzerland       & Deep Learning                     & To automatically estimate a patient's energy and macronutrient intake by analyzing images of their meals before and after consumption.                                                                                                 \\
    \addlinespace
    \citeauthor{petrauskasExplainableArtificialIntelligencebased2021}, \citeyear{petrauskasExplainableArtificialIntelligencebased2021} & Lithuania         & Knowledge-Based \& Rule-Based     & To assess nutrition-related factors (symptoms) and to determine the likelihood of geriatric patient health risks associated with four syndromes: malnutrition, oropharyngeal dysphagia, dehydration, and eating disorders in dementia. \\
    \addlinespace
    \citeauthor{pfistererEnhancingFoodIntake2022}, \citeyear{pfistererEnhancingFoodIntake2022}                                         & Canada            & Deep Learning                     & To provide an automated, accurate, and objective method for tracking food and nutrient intake in long-term care residents to help prevent and manage malnutrition.                                                                     \\
    \addlinespace
    \citeauthor{rameshNovelRulebasedRecommender2021}, \citeyear{rameshNovelRulebasedRecommender2021}                                   & India             & Machine Learning (General)        & To propose a plausible model to suggest food items that are applicable for the Indian elderly diabetic population based on medical test results and personal preferences.                                                              \\
    \addlinespace
    \citeauthor{seintMedicationMealIntake2018}, \citeyear{seintMedicationMealIntake2018}                                               & Japan             & Machine Learning (General)        & To establish a video monitoring system for the act of taking medication and eating activity by using human-object interaction.                                                                                                         \\
    \addlinespace
    \citeauthor{seokFeasibilityEfficacyTouchCare2022}, \citeyear{seokFeasibilityEfficacyTouchCare2022}                                 & Republic of Korea & Machine Learning (General)        & To indirectly monitor eating and cooking behaviors via sensor data and provide AI-driven feedback to improve health-related habits and nutritional status.                                                                             \\
    \addlinespace
    \citeauthor{spoorenImprovingRecommendationsMeals2023}, \citeyear{spoorenImprovingRecommendationsMeals2023}                         & The Netherlands   & Machine Learning (General)        & To personalize meal and product recommendations to reduce user clicks in a diet tracking app for older adults.                                                                                                                         \\
    \addlinespace
    \citeauthor{suryawanshiMalnutritionIdentificationGeriatric2021}, \citeyear{suryawanshiMalnutritionIdentificationGeriatric2021}     & India             & Machine Learning (General)        & To design a system to identify the nutritional status of geriatric patients, classify malnutrition risk, and identify patients who will benefit from nutrition therapy using data mining on screening form parameters.                 \\
    \addlinespace
    \citeauthor{wangDetectionEatingGestures2024}, \citeyear{wangDetectionEatingGestures2024}                                           & Belgium           & Deep Learning                     & To develop an automated system to detect and segment the time interval of eating gestures in older adults using wearable IMU sensors.                                                                                                  \\
    \addlinespace
    \citeauthor{wangAreNutritionalLabeling2024}, \citeyear{wangAreNutritionalLabeling2024}                                             & China             & Natural Language Processing (NLP) & To assist older adults with food identification and dietary planning for better dietary management.                                                                                                                                    \\
\end{longtable}
