\clearpage
\section{Informed Consent Form}
\label{app:consent_form}

\textbf{Study Title:} User Evaluation of "Snap and Say": An AI Dietary Assistant for Older Adults \\
\textbf{Principal Investigator:} Fabian W. N. Schreuder \\
\textbf{Institution:} Information Technology Group, Wageningen University \& Research

\subsection*{1. What is this study about?}

You are invited to participate in a research study evaluating \textbf{"Snap and Say,"} a new mobile web application designed to help older adults track their nutrition. The purpose of this study is to determine if using Artificial Intelligence (AI) and voice commands can make it easier to log meals compared to typing them out by hand.

\subsection*{2. What will I be asked to do?}

If you agree to participate, you will be asked to complete a short session lasting approximately \textbf{15--20 minutes}. During this session, you will:

\begin{enumerate}
    \item \textbf{Photograph a Meal:} Use the app to take a picture of a simulated meal provided by the researcher.
    \item \textbf{Interact with the Assistant:} If the AI is unsure about an ingredient, it may ask you a simple question (e.g., "Is the dressing Ranch or Italian?"). You will answer using your voice.
    \item \textbf{Complete a Survey:} Fill out a brief questionnaire about your experience, focusing on how easy or difficult the task felt.
\end{enumerate}

\subsection*{3. What data will be collected?}

To evaluate the system, the application will record and store:

\begin{itemize}
    \item \textbf{Images:} The photos of the food you upload.
    \item \textbf{Audio:} Your voice recordings when answering the assistant's questions.
    \item \textbf{Logs:} The time it took to complete the task and the accuracy of the nutritional estimate.
\end{itemize}

\subsection*{4. How will my privacy be protected?}

Your privacy is our priority.

\begin{itemize}
    \item \textbf{Anonymity:} Your personal identity will \textbf{not} be linked to your data. You will be assigned an \texttt{anonymous\_id} (a random code) for the duration of the study.
    \item \textbf{Data Security:} All data, including images and voice transcripts, are stored securely in an encrypted database (Supabase/PostgreSQL).
    \item \textbf{Usage:} The data will be used solely for scientific analysis and publication. No raw audio or photos identifying you personally will be shared publicly without explicit separate consent.
\end{itemize}

\subsection*{5. Risks and Benefits}

\begin{itemize}
    \item \textbf{Risks:} There are no known physical risks associated with this study. You may experience minor fatigue or frustration if technical issues arise, similar to using a standard smartphone app.
    \item \textbf{Benefits:} While there is no direct monetary benefit to you, your participation will help researchers design better health technology for older adults.
\end{itemize}

\subsection*{6. Voluntary Participation}

Participation is entirely voluntary. You may refuse to answer any question or stop the experiment at any time without providing a reason.

\vspace{1em}
\hrule
\vspace{1em}

\subsection*{7. Contact Information}

If you have questions about this research, please contact the lead researcher:

\begin{itemize}
    \item \textbf{Name:} Fabian W. N. Schreuder
    \item \textbf{Email:} fabian.schreuder@wur.nl
    \item \textbf{Department:} Information Technology Group, Wageningen University \& Research
\end{itemize}

\vspace{1em}
\hrule
\vspace{1em}

\subsection*{8. Participant Declaration}

By signing below, I acknowledge that:

\begin{itemize}
    \item I have read and understood the information provided above.
    \item I have had the opportunity to ask questions.
    \item I understand that my voice and images will be processed by the system.
    \item I agree to participate voluntarily.
\end{itemize}

\vspace{2em}

\textbf{Participant Name (Printed):} \rule{8cm}{0.4pt}

\vspace{1em}

\textbf{Participant Signature:} \rule{8cm}{0.4pt}

\vspace{1em}

\textbf{Date:} \rule{4cm}{0.4pt}

\vspace{1em}
\hrule
\vspace{1em}

\subsection*{Researcher Use Only}

\textbf{Participant ID:} \rule{6cm}{0.4pt} \\
\textbf{Date of Session:} \rule{6cm}{0.4pt}
