\clearpage
\section{User Feedback Survey}

\begin{center}
    {\Huge \textbf{User Feedback Survey}} \\[0.5em]
    {\Large "Snap and Say" Study}
\end{center}

\vspace{0.5cm}

\noindent
\begin{tabular}{@{}ll@{}}
    \textbf{Participant ID:} & \underline{\hspace{5cm}} \\[1em]
    \textbf{Date:}           & \underline{\hspace{5cm}}
\end{tabular}

\vspace{1cm}
\hrule
\vspace{0.5cm}

\noindent \textit{Instructions: Please circle the number that best describes your experience today. There are no right or wrong answers.}

\section*{Part 1: How much effort was it?}
\textit{(Based on the tasks you just finished)}

\vspace{1em}

\noindent \textbf{1. Mental Effort} \\
How much thinking was required to decide when to speak vs. when to click "Upload"?
\surveyrating{Very Low}{Very High}

\noindent \textbf{2. Physical Effort} \\
How hard was it to hold the phone, aim the camera, or tap the buttons?
\surveyrating{Very Easy}{Very Hard}

\noindent \textbf{3. Waiting Time} \\
How bothersome was the time you spent waiting for the assistant to think?
\surveyrating{Not Bothersome}{Very Bothersome}

\noindent \textbf{4. Frustration Level} \\
How discouraged, irritated, or annoyed did you feel during the task?
\surveyrating{Not Frustrated}{Very Frustrated}

\noindent \textbf{5. Success} \\
How successful did you feel in capturing your meal correctly?
\surveyrating{Very Successful}{Unsuccessful}

\clearpage

\section*{Part 2: Working with the Assistant}
\textit{(Thinking specifically about the Voice Screen)}

\vspace{1em}

\noindent \textbf{6. Helpfulness of Questions} \\
When the Assistant asked you about the salad dressing, did that feel helpful or annoying?
\surveyratingmid{Annoying}{Okay}{Helpful}

\noindent \textbf{7. Understanding the Question} \\
Was it clear \textit{why} the Assistant was asking you a question?
\surveyrating{Confusing}{Very Clear}

\noindent \textbf{8. Voice Control} \\
Was speaking to the app easier or harder than typing?
\surveyratingmid{Much Harder}{Same}{Much Easier}

\noindent \textbf{9. Vibration, Sound, \& Visual Feedback} \\
Did the vibration, sound, and visual wave help you know the app was listening?
\surveyrating{Not Enough}{Very Clear}

\vspace{0.5cm}
\hrule
\vspace{0.5cm}

\section*{Part 3: Overall Impressions}

\vspace{1em}

\noindent \textbf{10. I think I would like to use this system frequently.}
\surveyrating{Strongly Disagree}{Strongly Agree}

\noindent \textbf{11. I found the system unnecessarily complex.}
\surveyrating{Strongly Disagree}{Strongly Agree}

\noindent \textbf{12. I felt very confident using the system.}
\surveyrating{Strongly Disagree}{Strongly Agree}

\noindent \textbf{13. I imagine that most people would learn to use this system very quickly.}
\surveyrating{Strongly Disagree}{Strongly Agree}

\vspace{0.5cm}
\hrule
\vspace{0.5cm}

\section*{Part 4: Open Feedback}

\vspace{1em}

\noindent \textbf{14. What was the hardest part of using the app?} \\[2em]
\rule{\textwidth}{0.4pt} \\[1cm]
\rule{\textwidth}{0.4pt}

\vspace{1cm}

\noindent \textbf{15. If you could change one thing, what would it be?} \\[2em]
\rule{\textwidth}{0.4pt} \\[1cm]
\rule{\textwidth}{0.4pt}
