\clearpage
\section{Participant Task Script}
\label{app:task_script}

This script guides the participant through \textbf{two specific scenarios} designed to test the full range of the system:

\begin{enumerate}
    \item \textbf{Task A (The Simple Meal):} Validates the "Single-Shot" low-friction flow.
    \item \textbf{Task B (The Complex Meal):} Validates the "Agentic Logic" and voice interaction flow.
\end{enumerate}

\subsection*{System Setup \& Pre-Session Checklist}
\begin{enumerate}
    \item \textbf{Battery Check:} Ensure the smartphone is $>$50\% charge (camera and AI processing drain battery quickly).
    \item \textbf{Audio/Visual Check:} Ensure screen brightness is at 100\% so the visual wave is clearly visible against glare (as audio feedback is minimal).
    \item \textbf{Network:} Confirm the device is on a low-latency Wi-Fi network (essential for the "Agentic Logic" to respond quickly).
    \item \textbf{Screen Setup:} Ensure the laptop/monitor is at eye level to simulate a dining experience, not flat on a table (which increases glare).
    \item \textbf{Image Readiness:} Ensure the two meal images are ready to display on a laptop/tablet monitor.
\end{enumerate}
\vspace{1em}
\hrule
\vspace{1em}

\subsection*{Participant Task Sheet: The "Digital Buffet" Session}

\noindent\fbox{\parbox{\textwidth}{\textbf{READ ALOUD TO PARTICIPANT:}

        "Welcome to the session. We are testing a new food logging app called 'Snap and Say'.

        \textbf{Three quick rules:}
        \begin{enumerate}
            \item \textbf{I am not the designer.} You won't hurt my feelings if you say something is ugly or confusing. Please be brutally honest.
            \item \textbf{Think Aloud.} As you use the app, please try to tell me what you are looking at. For example: 'I'm looking for the save button' or 'I'm wondering if it heard me.'
            \item \textbf{This is a test of the app, not you.} If you get stuck, it helps us improve the software.
        \end{enumerate}

        Do you have any questions before we look at the first meal?"
    }
}

\vspace{1em}
\hrule
\vspace{1em}

\subsection*{Task A: The Simple Meal (Practice)}

\textit{Please look at the computer screen. It currently shows a plate of \textbf{Roasted Chicken and Vegetables}.}

\subsubsection*{1. Open the App}

\begin{itemize}
    \item Take the smartphone and tap the \textbf{"New Meal"} button.
\end{itemize}

\subsubsection*{2. Capture the "Meal"}

\begin{itemize}
    \item Hold the phone about \textbf{30 cm (12 inches)} away from the computer screen.
    \item \textit{Tip:} If the screen looks too bright or has wavy lines, move the phone back slightly.
    \item Tap the \textbf{Camera Button} to take the picture.
    \item Tap \textbf{"Use Photo"} to confirm the image preview.
\end{itemize}

\subsubsection*{3. The Voice Screen (Required Step)}

\begin{itemize}
    \item The app will now show the \textbf{Voice Recording} screen.
    \item \textbf{Note:} You must pass through this screen for every meal.
    \item Since this meal is simple and needs no explanation, simply tap the \textbf{"Upload"} button (without recording) to proceed.
\end{itemize}

\subsubsection*{4. Review the Result}

\begin{itemize}
    \item Wait a few seconds for the assistant to think.
    \item Because this meal is easy to see, the assistant should show you the list (Chicken, Carrots, Potatoes) immediately.
    \item Tap the green \textbf{"Save Log"} button to finish.
\end{itemize}

\textit{(Great job! Now we will try a harder one.)}

\vspace{1em}
\hrule
\vspace{1em}

\subsection*{Task B: The Complex Meal (The Real Test)}

\textit{The researcher will now change the image on the screen. It is now a \textbf{Mixed Green Salad}.}

\subsubsection*{1. Capture and Upload}

\begin{itemize}
    \item Tap \textbf{"New Meal"} again.
    \item Point the camera at the salad on the screen.
    \item Tap the \textbf{Camera Button} and then \textbf{"Use Photo"}.
    \item When the Voice Screen appears, tap \textbf{"Upload"} immediately to let the AI analyze the photo first.
\end{itemize}

\subsubsection*{2. The Assistant Needs Help}

\begin{itemize}
    \item \textit{Please wait.} This time, the assistant might be unsure because it cannot see if there is dressing or oil hidden in the salad.
    \item \textbf{Watch the screen:} The assistant will pause and ask you a question (e.g., \textit{"Is there any dressing on this?"}).
\end{itemize}

\subsubsection*{3. Answer with Your Voice}

\begin{itemize}
    \item We designed this app so you don't have to type.
    \item \textbf{Step 1:} Tap the \textbf{Microphone Button ONCE} to start listening.
    \item \textbf{Step 2:} Speak clearly to the phone: \textit{"Yes, it has Italian dressing."}
    \item \textbf{Step 3:} Tap the \textbf{Microphone Button AGAIN} to stop recording.
    \item The wave on the screen will stop moving.
\end{itemize}

\subsubsection*{4. Finish}

\begin{itemize}
    \item The assistant will update the list to include the dressing.
    \item Tap \textbf{"Save Log"}.
\end{itemize}

\vspace{1em}
\hrule
\vspace{1em}

\subsection*{End of Session}

Thank you! Please hand the device back to the researcher.
We will now ask you a few short questions about how easy or difficult that felt.
\vspace{1em}
\hrule
\vspace{1em}

\section*{Researcher Protocol: "Digital Buffet" Checklist}

\textbf{Setup for Task B (Complex):}

\begin{itemize}
    \item \textbf{Image Selection:} Display a salad image specifically chosen from the "Complex" class (e.g., >3 ingredients, hidden density).
    \item \textbf{System Check:} Ensure the backend confidence threshold is active so the question generation node is triggered.
\end{itemize}

\textbf{Critical Observations (Mark Yes/No):}

\begin{itemize}
    \item \textbf{Glare Management:} Did the user struggle to get a clear photo of the screen?
    \item \textbf{Flow Confusion:} In Task A, did the user understand they had to tap "Upload" to skip the voice recording?
    \item \textbf{Motor Control:} Did the \textbf{"Tap-to-Toggle"} (tap once) work intuitively, or did they try to hold the button down?
\end{itemize}

\textbf{Next Step:} Administer the Survey.